Decentralized state machine using Nakamoto (probabilistic) consensus\\\medskip Research Project 2022

### 2022-11-01 article/article.tex
more article

\documentclass[11pt, twocolumn]{article}
\usepackage[T1]{fontenc}
\usepackage{amsmath, amssymb}
\usepackage{mathtools}
\usepackage{graphicx}
\usepackage[utf8]{inputenc}
\usepackage{fdsymbol}
\usepackage{textgreek}
\usepackage{natbib}
\usepackage{hyperref}
\usepackage{nameref}
\usepackage{url}
\usepackage{array}
\usepackage{csquotes}
\usepackage{caption}
\usepackage{algpseudocode}
\usepackage[toc,page]{appendix}
\DeclarePairedDelimiter\floor{\lfloor}{\rfloor}
\providecommand{\myceil}[1]{\left \lceil #1 \right \rceil }
\hypersetup{
    colorlinks,
    citecolor=black,
    filecolor=black,
    linkcolor=black,
    urlcolor=black
}

@@

In snowball, each node has a counter of "confidence" which can tip towards one way or another. In my implementation, there is no binary consensus because the only value the system has is the log request.
When receiving an add-log request, the node follows this sequence :
\begin{algorithmic}
    \State from request (client or other node):
    \State $logToAdd$ (string)
    \State $logPosition$ (int)

    \State \\from node memory :
    \State $record$ (map[int]string)
    \State $cnt$ (map[int]int)
    \\
    \If{$record[logPosition]$ exists}
    \State reply $record[logPosition] == logToAdd$
    \Else
    \State $record[logPosition] \gets logToAdd$
    \State reply $True$
    \State initiateRequest($logToAdd$, $logPosition$)
    \EndIf

\end{algorithmic}

@@

3 parameters are used in my implementation :
\begin{itemize}
    \item sampleSize (k) : the number of nodes that are queried each round
    \item majorityThreshold (α) : the number of nodes that have to reply successfully to consider that the query is successfull.
          (with \begin{equation*} \alpha \geq \floor{k/2}  \end{equation*})
    \item counterThreshold (β) : the threshold of successfull requests above which the node stops initiating requests and only replies with its content. (for one position)
\end{itemize}

@@

\paragraph{β = 1}
With β = 1, the state is (c, reqs) with :
\begin{itemize}
    \item c = Number of colored nodes ($c \leq n$)
    \item reqs = Number of groups of requests which are queried for next round. (They all come from neutral nodes)
\end{itemize}
If the state converges to $c = n$, consensus is reached.
If the state converges to $reqs = 0$, with $c \neq 0$, consensus is not reached when the snowball run is done.

@@

Therefore, the probability of being at state (c, reqs) is (the term in the sum is considered 0 if it does not make sense, which is the case for some t) :
The sum represents a step of the resolution of one group of requests (t represent the number of neutral nodes touched)
\begin{equation*}
    P(c, reqs) = \sum_{t=0}^{k+1} {n-c+t \choose t}{c-t-1 \choose k-t}P(c-t, reqs+1-t)
\end{equation*}

\paragraph{β = 2}
With β = 2, the state is (c, s, rc, rs), with:
\begin{itemize}
    \item c = Number of colored nodes (cnt = 2)
    \item s = Number of semi-colored nodes (cnt = 1)
    \item rc = Number of groups of requests from colored nodes
    \item rs = Number of groups of requests from semi-colored nodes
\end{itemize}

@@

The term in the sums is considered 0 if it does not make sense.
First sum represents a step of the resolution of one group of requests, coming from a semi-colored node, with tc (ts) the number of colored (semi colored) nodes in the sample.
Second sum represents a step of the resolution of one group of requests, coming from a colored node. 
To do the calculation, we assume that requests from semi-colored nodes resolve faster.
\begin{multline*}
    P(c, s, rc, rs) = \\ \sum_{tv=0}^{k+1}\sum_{ts=0}^{k+1} {n-c-s+tv \choose tv}{s-tv-1+ts \choose ts}{c-ts \choose k-tv-ts}P(c-ts, s-tv+ts, rc-ts, rs+1-tv)
    \\ + \sum_{ts=0}^{k+1} {n-c-s+rs \choose tv}{s-rs+ts \choose ts}{c-ts-1 \choose k-rs-ts}P(c-ts, s-rs+ts, rc+1-ts, 0)
\end{multline*}
\\
With these formulas, we can compute the probability of not reaching consensus : view Appendices

@@

Number of requests
Number of requests
Number of requests
Number of requests
Number of requests
Number of requests
Number of requests
Number of requests
Number of requests
Number of requests
Number of requests
Number of requests

\subsection{In case of a failure}

Failure ?
Failure ?
Failure ?
Failure ?
Failure ?
Failure ?
Failure ?
Failure ?
Failure ?
Failure ?

@@

Conclusion
Conclusion
Conclusion
Conclusion
Conclusion
Conclusion
Conclusion
Conclusion
Conclusion

\onecolumn
\bibliographystyle{alpha}
\bibliography{biblio}

\begin{appendices}
    \label{Appendices}
    \begin{figure*}
        Probability of not reaching consensus (β = 1) : ε 
        \centering
        \includegraphics[width=16.5cm]{images/smallNoConsensusBetaOne.png}
    \end{figure*}
    \begin{figure*}
        Probability of not reaching consensus (β = 1) : -log(ε) 
        \centering
        \includegraphics[width=16.5cm]{images/smalllogNoConsensusBetaOne.png}
    \end{figure*}
    \begin{figure*}
        Probability of not reaching consensus (β = 2) : ε 
        \centering
        \includegraphics[width=16.5cm]{images/smallnoconsensus.png}
    \end{figure*}
    \begin{figure*}
        Probability of not reaching consensus (β = 2) : -log(ε) 
        \centering
        \includegraphics[width=16.5cm]{images/smalllognoconsensus.png}
    \end{figure*}
\end{appendices}
\end{document}

>>>>

\documentclass[11pt, twocolumn]{article}
\usepackage[T1]{fontenc}
\usepackage{amsmath, amssymb}
\usepackage{mathtools}
\usepackage{graphicx}
\usepackage[utf8]{inputenc}
\usepackage{fdsymbol}
\usepackage{textgreek}
\usepackage{natbib}
\usepackage{hyperref}
\usepackage{nameref}
\usepackage{url}
\usepackage{array}
\usepackage{csquotes}
\usepackage{caption}
\usepackage{algpseudocode}
\usepackage[a4paper, total={6.5in, 9in}]{geometry}
\usepackage[toc,page]{appendix}
\DeclarePairedDelimiter\floor{\lfloor}{\rfloor}
\providecommand{\myceil}[1]{\left \lceil #1 \right \rceil }
\hypersetup{
    colorlinks,
    citecolor=black,
    filecolor=black,
    linkcolor=black,
    urlcolor=black
}

@@

In snowball, each node has a counter of "confidence" which can tip towards one way or another. In my implementation, there is no binary consensus because the only value the system has is the log request.
When receiving an add-log request, the node follows this sequence, in which $logToAdd$($string$) and $logPosition$($int$) are variables from the request (client or other node) and $record$($map[int]string$) and $cnt$($map[int]int$) are variables from the node memory:
\begin{algorithmic}
    \If{$record[logPosition]$ exists}
    \State reply $record[logPosition] == logToAdd$
    \Else
    \State $record[logPosition] \gets logToAdd$
    \State reply $True$
    \State initiateRequest($logToAdd$, $logPosition$)
    \EndIf
\end{algorithmic}

@@

3 parameters are used in my implementation :
\begin{itemize}
    \item sampleSize (k) : the number of nodes that are queried each round
    \item majorityThreshold (α) : the number of nodes that have to reply successfully to consider that the query is successfull.
          (with $\alpha \geq \lceil k/2 \rceil$)
    \item counterThreshold (β) : the threshold of successfull requests above which the node stops initiating requests and only replies with its content. (for one position)
\end{itemize}

@@

\paragraph{β = 1}
With β = 1, the state is (c, reqs) with :
\begin{itemize}
    \item c = Number of colored (cnt = 1) nodes ($c \leq n$)
    \item reqs = Number of groups of requests which are queried for next round. (They all come from neutral nodes)
\end{itemize}
If the state converges to $c = n$, consensus is reached.
If the state converges to $reqs = 0$, with $c \neq 0$, consensus is not reached when the snowball run is done.

@@

Therefore, the probability of being at state (c, reqs) is (the term in the sum is considered 0 if it does not make sense, which is the case for some t) :
\begin{equation*}
    P(c, reqs) = \sum_{t=0}^{k+1} {n-c+t \choose t}{c-t-1 \choose k-t}P(c-t, reqs+1-t)
\end{equation*}
The sum represents the resolution of one group of requests of the consensus protocol for one position (t represent the number of neutral nodes touched)

\paragraph{β = 2}
With β = 2, the state is (c, s, rc, rs), with:
\begin{itemize}
    \item c = Number of colored (cnt = 2) nodes
    \item s = Number of semi-colored (cnt = 1) nodes
    \item rc = Number of groups of requests from colored nodes
    \item rs = Number of groups of requests from semi-colored nodes
\end{itemize}

@@

The term in the sums is considered 0 if it does not make sense.
First sum represents a step of the resolution of one group of requests, coming from a semi-colored node, with tc (ts) the number of colored (semi colored) nodes in the sample.
Second sum represents a step of the resolution of one group of requests, coming from a colored node.
To do the calculation, we assume that requests from semi-colored nodes resolve faster.
\begin{multline*}
    P(c, s, rc, rs) = \\ \sum_{tv=0}^{k+1}\sum_{ts=0}^{k+1} {n-c-s+tv \choose tv}{s-tv-1+ts \choose ts}{c-ts \choose k-tv-ts}P(c-ts, s-tv+ts, rc-ts, rs+1-tv)
    \\ + \sum_{ts=0}^{k+1} {n-c-s+rs \choose tv}{s-rs+ts \choose ts}{c-ts-1 \choose k-rs-ts}P(c-ts, s-rs+ts, rc+1-ts, 0)
\end{multline*}
\\
With these formulas, we can compute the probability of not reaching consensus : view Appendices
The scripts which were used to compute the graphs in Appendix are available at \href{https://github.com/Nicolascrd/researchProjectConsensus}{github.com/Nicolascrd/researchProjectConsensus}

@@

\paragraph{Raft (classical)}
When the leader is up and giving instruction to all the followers, followers don't have to initiate requests.
Therefore in normal functioning, the leader node will make $n-1$ requests every heartbeat.
In addition to this, every client request will trigger $n-1$ requests from the leader, to update followers. If the request was directed from the client to a follower, one request is initiated
by the follower to transfer the request to the leader.

\paragraph{Snowball (Nakamoto)}
The number of requests required to reach consensus on an add-log request is $k*\beta$ for each node.

\subsection{In case of a failure}
\paragraph{Fail-stop in Raft}
If a follower crashes, the leader stays in place. If updateSystem is activated, one request will be made to each follower to update its knowledge of the system. That represents $n-2$ requests in this case

If the leader crashes, nodes will switch from the follower status to the candidate status, because they don't receive the heartbeat.
However, the number of requests that will be made to elect a new leader depends on a number of factors.
With a few nodes, the process happens with only one request : the first node to tick (notice that the heartbeat was not received) switches to candidate status, triggers a round of vote, collects all the votes and establishes itself as the new leader.
In the meantime, the other nodes did not switch to candidate and start receiving the heartbeats from the new leader.
With many nodes, nodes might tick "at the same time" and therefore candidate at the same time which can trigger many requests. To avoid an infinite escalation of currentTerm, which is the value incremented by each candidate, the ticker is supposed to happen randomly in some time window.

Overall the Raft is very efficient at managing crash failures.

\paragraph{Fail-stop in Snowball}
If a node crashs in Snowball, some add-log request will not find a majority and therefore the consensus will be slower and require more requests.
However, in production, the Snowball state-machine would include a mechanism to update the network knowledge among members, which can take the form of a Snowball run, and therefore continue as if no node had failed.
Snowball is also very efficient at managing crash failures.

\paragraph{Byzantine failures}
The architecture required to test properly my implementations for byzantine failures goes beyong this project. But we can still think about the resilience of each implementation.
With Raft, a byzantine node can easily take the control on the network by becoming the leader. It only needs to start as a follower, choose a currentTerm bigger than the current one and ask for votes, which other nodes will give.
Once elected leader, the byzantine node can implement any information among followers. The state-machine is unusable and even potentially dangerously infected.

With Avalanche, to begin with, a node cannot easily change the information that is already in place among nodes. This would require the consensus to switch from one way to the other, which is nearly impossible. However, in my implementation, the client cannot either change a log at one position.
One thing the byzantine node cannot easily do in the Snowball implementation is register different informations among multiple nodes. Indeed, these nodes will want to verify the information by quering other nodes to reach consensus. The system will therefore go one way or the other.

@@

We demonstrated that it was possible to build a decentralized state machine, that could be based on a Nakamoto consensus algorithm and had interesting properties.
However, more research is required to use this approach at a greater scale.
In particular, it would be interesting to approximate the value of ε $\forall n, k, \beta$. 
Also, a comparison focused on Byzantine Fault Tolerance would be interesting. To do that one would need to use PBFT \cite{pbft} instead of the Raft.

\onecolumn

\begin{appendices}
    \bibliographystyle{alpha}
    \bibliography{biblio}
    \label{Appendices}
    \begin{figure*}
        Probability of not reaching consensus (β = 1) : ε
        \centering
        \includegraphics[width=16.5cm]{images/smallNoConsensusBetaOne.png}
    \end{figure*}
    \begin{figure*}
        Probability of not reaching consensus (β = 1) : -log(ε)
        \centering
        \includegraphics[width=16.5cm]{images/smalllogNoConsensusBetaOne.png}
    \end{figure*}
    \begin{figure*}
        Probability of not reaching consensus (β = 2) : ε
        \centering
        \includegraphics[width=16.5cm]{images/smallnoconsensus.png}
    \end{figure*}
    \begin{figure*}
        Probability of not reaching consensus (β = 2) : -log(ε)
        \centering
        \includegraphics[width=16.5cm]{images/smalllognoconsensus.png}
    \end{figure*}
\end{appendices}
\end{document}
